---
title: | 
  | \vspace{-1.8cm} MSSC 5780 Homework 4
output: 
  pdf_document:
    number_sections: true
fontsize: 11pt
geometry: margin=1in
documentclass: article
---

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      var: "{\\mathrm{Var}}",
      cov: "{\\mathrm{Cov}}",
      bsbeta: "{\\boldsymbol{\\beta}}",
      bsalpha: "{\\boldsymbol{\\alpha}}",
      bsep: "{\\boldsymbol{\\epsilon}}"
    }
  }
});
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
issue_due_date <- format(seq.Date( Sys.Date(), length=2, by='7 days'),  '%B %d %Y')
issue_due_date <- paste(issue_due_date, "11:59 PM")
```



\begin{center}
\textbf{Due Date: November 4, 2022 11:59 PM}
\textbf{Henrique Medeiros Dos Reis}
\end{center}

# Homework Instruction and Requirement
- Homework 4 covers course materials of Week 1 to 9.

- Please submit your work in **one** **PDF** file including all parts to **D2L > Assessments > Dropbox**. *Multiple files or a file that is not in pdf format are not allowed.*

- In your homework, please number and answer questions **in order**. 

- Your answers may be handwritten on the Mathematical Derivation and Reasoning part. However, you need to scan your paper and make it a PDF file. 

- Your entire work on Statistical Computing and Data Analysis should be completed by any word processing software (Microsoft Word, Google Docs, [(R)Markdown](https://rmarkdown.rstudio.com/), [LaTex](https://www.latex-project.org/), etc) and your preferred programming language. Your document should be a PDF file.

- Questions starting with **(MSSC)** are for MSSC 5780 students.

- It is your responsibility to let me understand what you try to show. If you type your answers, make sure there are no typos. I grade your work based on *what you show, not what you want to show*. If you handwrite your answers, write them neatly. If I can’t read your sloppy handwriting, your answer is judged as wrong.



# Mathematical Derivation and Reasoning


<!-- ## Math Derivation -->
The simple linear regression and multiple linear regression models and notations are the same as defined in our course slides and textbook.



1. **(Optional)** Suppose that we have fit the least-squares regression model $\hat{y} = b_0 + b_1x_1$ but the response is affected by a second variable $x_2$ such that the true regression function is $$E(y) = \beta_0 + \beta_1x_1 + \beta_2x_2$$
The coefficient $b_1$ in the original simple linear regression model is no longer unbiased. Show the bias in $b_1$, i.e., $E(b_1) - \beta_1$.


Let the hat matrix ${\bf H}_{n \times n} = {\bf X(X'X)}^{-1}{\bf X'}$ from a multiple linear regression model.


2. **(Optional)** Let $h_{ii}$ be the $i$-th diagonal element of ${\bf H}$. Show that for the multiple linear regression model $\frac{1}{n} \le h_{ii} \le 1$.
<!-- 5. **(MSSC)** Show that for all nonzero vectors ${\bf y}$, ${\bf y}'{\bf A} {\bf y} \ge 0$. -->

3. **(Optional)** Show that the weighted least squares estimator minimizes $$({\bf y} - {\bf X} \boldsymbol{\beta})'{\bf W}({\bf y} - {\bf X}\boldsymbol{\beta})$$ where ${\bf W}$ is the diagonal matrix whose diagonal elements are weights $w_1, w_2, \dots, x_n$.



# Statistical Computing and Data Analysis


Please perform a data analysis using $\texttt{R}$ or your preferred language. **Any results should be generated by computer outputs, and your work should be done entirely by a computer. Handwriting is not allowed. Relevant code should be attached.**

## Diagnostics on Gasoline Mileage Data
We use the same data set `mpg.csv` for data analysis. Consider the multiple regression model $y = \beta_0 + \beta_1x_1 + \beta_6x_6+\epsilon$ fit to the gasoline mileage data in your Homework 3.

```{r}
mpgData<-mpgData <- read.csv("mpg.csv")
mpg_engDisp_carBar_lm <- lm(y~x1+x6, data = mpgData)
mpg_engDisp_carBar_lm
```

1. Compare R-student residuals $t_i$ with Student-t $t_{n-p-1}$ using a qqplot. Generate the histogram and density plot of $t_i$ as well. Does there seem to be any problem with the normality assumption?
 
\quad\quad\ **Solution: **
```{r}
car::qqPlot(mpg_engDisp_carBar_lm, id = TRUE, col.lines = "blue", 
            reps = 1000, ylab = "Ordered R-Student Residuals", pch = 16)
```
```{r}
r_stud <- rstudent(mpg_engDisp_carBar_lm)

hist(r_stud, prob = TRUE, breaks = 10, xlab = "R-Student Residuals", main = "")
lines(density(r_stud, adjust = 2), col = 4, lwd = 2)
```

We can see that the qq-plot seems fine, but the histogram seems to have heavier tails, which indicates that this is not exactly normal. And because this has heavier tails, it probably is a t-dist.



2. Perform the Box-Cox method and discuss the necessity of any transformation on $y$.

\quad\quad\ **Solution: **
```{r}
carPT<-car::powerTransform(mpg_engDisp_carBar_lm, family = "bcPower")
summary(carPT)
```

Looking at the summary enough evidence that the transformation at $\lambda$ = -0.1483 should be done. 


3. Use $\lambda = 0$ (log transformation) and the $\lambda$ selected by the Box-Cox method to refit the model to the transformed data. Compare their R-student residuals with the R-student residuals from the non-transformed data using a boxplot.

\quad\quad\ **Solution: **
```{r}
log_mpg_engD_carB <- lm(log(y) ~ x1 + x6, data = mpgData)
lambda_mpg_engD_carB <-lm(y^(-0.1483) ~ x1+x6, data = mpgData)

mat <- matrix(r_stud)
for (lam in c(-0.1483, 0)) {
    refit <- update(
        mpg_engDisp_carBar_lm, car::bcPower(y, lam) ~ .
        )
    mat <- cbind(rstudent(refit), mat)
}
colnames(mat) <- c(-0.1483, "log", 1)
boxplot(
    mat, id = FALSE, 
    xlab = expression("Powers," ~ lambda),
    ylab = expression(
        "R-Student Residuals for " 
        ~ y ^ (lambda))
    )
```
We can see that the transformations for log and -0.14 box plot are the same, while the non-transformed one is not symmetric.

4. Construct a plot of the R-student residuals $t_i$ versus the fitted responses and the Tukey's spread-level plot. Any sign of violation of constant variance?

\quad\quad\ **Solution: **
```{r}
plot(log_mpg_engD_carB$fitted.values, log_mpg_engD_carB$residuals)
abline(0,0, col="red")

car::spreadLevelPlot(mpg_engDisp_carBar_lm, smooth = FALSE)
```
It is possible to see that there definitely is some sign of constant variance.
Since the line, in the plot with the blue line, has a positive slope, the variance is not constant. 

5. In fact, we can perform some formal hypothesis testing on constant variance like $H_0:$ Constant variance vs. $H_1:$ Variance changes with $E(y\mid x)$. Use `car::ncvTest()` to perform the test. Explain the testing result.

\quad\quad\ **Solution: **
```{r}
car::ncvTest(mpg_engDisp_carBar_lm)
```
Since the p-value is $0.046782$ which is smaller than $\alpha = 0.05$. Which leads to reject null hypothesis. Therefore, not constant variance.



## Influence Diagnostics on Squid Data
An experiment was conducted to study the size of squid eaten by sharks and tuna. The regressors are characteristics of the beak or month of the squid. The `squid.csv` data contain the variables

+ $x_1$: Rostral length in inches
+ $x_2$: Wing length in inches
+ $x_3$: Rostral to notch length
+ $x_4$: Notch to wing length
+ $x_5$: Width in inches
+ $y$: Weights in pounds
```{r}
squidData <- read.csv("squid.csv")
squidReg <- lm(y ~ x1+x2+x3+x4+x5, data = squidData)
```


Perform a thorough leverage and influence diagnostics of the squid data.

1. Compute R-studentized residuals, hat values, Cook's distance, DFFITS, DFBETAS, and COVRATIO measures. Describe how you detect leverage and influential points. Discuss the effect of data points on coefficients, fitted values, and precision of coefficients. (The `influence.measures()` function provides all influence measures.)

\quad\quad\ **Solution: **
```{r}
summary(influence.measures(squidReg))
```
Looking at this, we can see that the points 2, 18, and 19 got flagged as influential points. The reason for this is because in at least one of the measures, this points got a value that is being influential to that specific measurement. For example, looking at df betas 2, 2 and 18 are influential points. And looking at cov.r, point 19 is influential. 

2. Let's use some visualization tools. 
    * Create the bubble plot.
    * Create the influence index plot (`car::influenceIndexPlot()`)
    
\quad\quad\ **Solution: **

```{r}
car::influencePlot(squidReg)
car::influenceIndexPlot(squidReg)
```
And now looking at the graphs, it agrees with the explanation in number 2. 

3. Produce the added-valued plot (`car::avPlots()`) for each regressor $x_i, i = 1, \dots, 5$. Are there any joint influence of data points on the regression coefficients?

\quad\quad\ **Solution: **
```{r}
car::avPlots(squidReg)
```
Looking at these, we can see that points 2 and 18 are jointly influential, almost in all of the the coefficients. they "put down" the line. 


4. **(MSSC)** Numerically verify the following properties of the added-valued plot.
    (a) The slope of the least squares *simple* regression line of $e(y \mid x_{(1)})$ on $e(x_1 \mid x_{(1)})$ is the same as the least-squares slope $b_1$ for $x_1$ in the full *multiple* regression.
    (b) The residuals from the simple regression $e(y \mid x_{(1)})$ vs. $e(x_1 \mid x_{(1)})$ are the same as the residuals $e_i$ from the full multiple regression.
    (c) The standard error of $b_1$ is $s / \sqrt{\sum_{i=1}^ne_i^2(x_1 \mid x_{(1)})}$, where $s = \sqrt{MS_{res}}$ from the full multiple regression.

\quad\quad\ **Solution: **

\quad\quad\ **a-)**

```{r, include=FALSE}
#a-)
my_coef <- summary(squidReg)$coefficients;
y_hat <-lm(y~x2+x3+x4+x5, data=squidData);

e_i_yx <- squidData$y - y_hat$fitted.values;

x_1_reg <- lm(x1 ~ x2+x3+x4+x5, data = squidData);
#x_1_new_vals <- x_1_reg$fitted.values
e_i_xx <- squidData$x1 - x_1_reg$fitted.values;
```
```{r}
(slope <- summary(lm(e_i_yx~e_i_xx))$coefficients[2,1])
my_coef[2,1]
```

We  can see that they are the same

```{r}
#b-)
round(squidReg$residuals) == round(lm(e_i_yx~e_i_xx)$residuals)
```
As we  can see, they are the same

```{r}
s<- sqrt((sum(squidReg$residuals^2))/(22-5-1))
(b1<- s/sqrt(sum(e_i_xx^2)))
sum_squid<-summary(squidReg)
sum_squid$coefficients[ 2, 2]
```
Once again, they are the same. 

## Simulation
Consider a simple linear regression model $y_i =\beta_0 +  \beta_1x_i + \epsilon_i,   i = 1, 2, \dots, n$ with $\epsilon_i \stackrel{\rm iid}{\sim} N(0, \sigma^2)$. Set $x_i = i$, $\beta_0 = 2$, $\beta_1 = -1.5$, $\sigma = 1.2$, $n= 15$.

1. **(MSSC)** Generate a sample $\{y_i\}_{i=1}^n$ from this model, and compute the regression coefficients $b_0$, $b_1$ and variance estimate $s^2$. Repeat 10,000 times.

\quad\quad\ **Solution: **

```{r}
#Residual standard error: 1.1 == s^2
beta_0 <- 2
beta_1 <- -1.5
n <- 15
sigm <- 1.2

x<- seq(1,15)

b_0<-c()
b_1<-c()
sigmas<-c()
for(i in 1:10000)
  {
    epi <- rnorm(15, 0, sigm)
    y <- beta_0+beta_1*x + epi
    my_lm<-lm(y~x)
    coeff <- summary(my_lm)$coefficients
    b_0<-c(b_0, coeff[1,1])
    b_1<-c(b_1, coeff[2,1])
    sigmas<-c(sigmas, (summary(my_lm))$sigma^2)
}
```


2. **(MSSC)** Compute the sample mean of $b_0$, $b_1$, and $s^2$. Compare them with their true expected value. What is your conclusion?

\quad\quad\ **Solution: **

```{r}
mean(b_0)
mean(b_1)
sqrt(mean(sigmas)) ###sigma squared
```
It is possible to see that the mean of the values for $b_0, b_1, and \sigma$ are almost equal to the values we set up at the beggining of the experiment, and are not exatly the same just because we added some noise to it. 

3. **(MSSC)** Plot the histograms of $b_0$, $b_1$, and $s^2$. Compare the histograms with their sampling distribution.

\quad\quad\ **Solution: **
```{r}
hist(b_0)
hist(b_1)
hist(sigmas)
```

The histograms for $b_0$ and $b_1$ seem to be normal distributed. But the $\sigma$ histogram is skewed to the right. Which seems more like a Gamma distribution. 

4. **(MSSC)** Compute the covariance matrix of $b_0$ and $b_1$ and the true covariance matrix.  Compare the two.

\quad\quad\ **Solution: **


```{r}
new_x <- cbind(1, x)
(cov_matrix<- cov(data.frame( b_0, b_1)))

sigm^2*solve(t(new_x)%*%new_x)

```
Once again, the values are almost the same, which only happens because we added some noise to the data. 

5. **(MSSC)** Find the variance of $s^2$. Use the fact that $\frac{(n-2)s^2}{\sigma^2} \sim \chi^2_{n-2}$ and $\mathrm{Var}(\chi^2_k) = 2k$ to get the true variance of $s^2$. Compare the two.

\quad\quad\ **Solution: **

Assuming normality, to get the the true variance of $s^2$, we find $Var(s^2)$. Since $Var(\frac{(n-2)s^2}{\sigma^2}) = Var(\chi^2) = 2(n-2)$, then, we have $Var(s^2) = \frac{(2\sigma^4)}{(n - 2)}$
```{r}
#sample
var(sigmas)

#true 
2*(sigm^4)/(n-2)

```
The values are almost the same. 

6. **(MSSC)** Suppose now $\beta_1 = 0.0001$. Generate the simulated data $\{x_i, y_i\}_{i=1}^n$ with $n = 10, 100, 1000, 10000$. Show that as $n$ increases, for the test $H_0: \beta_1 = 0$, its $t_{test}$ gets larger and its $p$-value gets smaller, and therefore $H_0$ is rejected for a sufficiently large $n$, even though $\beta_1$ is practically zero.


```{r}
#Residual standard error: 1.1 == s^2
beta_0 <- 2
beta_1 <- 0.0001
n_1 <- c(10,100,1000,10000)
p_vals<-c()
t_test<-c()
for(i in n_1)
{
  epi <- rnorm(i, 0, sigm)
  x<-seq(1,i)
  y<-beta_0+beta_1*x+epi
  
  p_vals<-c(p_vals,(t.test(beta_0+beta_1*x+epi, data= data.frame(x,y))$p.value))
  t_test<-c(t_test,(t.test(beta_0+beta_1*x+epi, data= data.frame(x,y))$statistic))
}
p_vals
as.numeric(t_test)
```

It is possible to see that the vector containing the p-values is definitely decreasing, and when $n=10000$, the p-value is just 0. And basically the same thing happens with T-statistic, except the opposite, it increases from $4.25$ when $n=10$ to $201.87$ when $n=10000$