---
title: | 
  | \vspace{-1.8cm} MATH 4780 (MSSC 5780) Homework 6
output: 
  pdf_document:
    number_sections: true
fontsize: 11pt
geometry: margin=1in
documentclass: article
---

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      var: "{\\mathrm{Var}}",
      cov: "{\\mathrm{Cov}}",
      bsbeta: "{\\boldsymbol{\\beta}}",
      bsalpha: "{\\boldsymbol{\\alpha}}",
      bsep: "{\\boldsymbol{\\epsilon}}"
    }
  }
});
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
issue_due_date <- format(seq.Date( Sys.Date(), length=2, by='7 days'),  '%B %d %Y')
issue_due_date <- paste(issue_due_date, "11:59 PM")
```

<!-- \begin{center} -->
<!-- __Due Date: `r issue_due_date[2]`__ -->
<!-- \end{center} -->
<!-- \center  \center -->
\begin{center}
\textbf{Due Date: November 4, 2022 11:59 PM}
\textbf{Henrique Medeiros Dos Reis}
\end{center}
<!-- <p style="text-align: center;">**Due Date: `r issue_due_date[2]`** </p> -->

<!-- \fontfamily{cmr} -->
<!-- \fontsize{12}{22} -->
<!-- \fontseries{b} -->
<!-- \selectfont -->

# Homework Instruction and Requirement
- Homework 6 covers course materials of Week 1 to 12.

- Please submit your work in **one** **PDF** file including all parts to **D2L > Assessments > Dropbox**. *Multiple files or a file that is not in pdf format are not allowed.*

- In your homework, please number and answer questions **in order**. 

- Your entire work should be completed by any word processing software (Microsoft Word, Google Docs, [(R)Markdown](https://rmarkdown.rstudio.com/), [LaTex](https://www.latex-project.org/), etc) and your preferred programming language. Your document should be a PDF file.

- It is your responsibility to let me understand what you try to show. If you type your answers, make sure there are no typos. I grade your work based on *what you show, not what you want to show*. If you choose to handwrite your answers, write them neatly. If I can’t read your sloppy handwriting, your answer is judged as wrong.


# Reading and Writing
In this course, we have been using the classical or frequentist approach to do a variety of statistical inferences, marginal $t$ test for $\beta_j$ and $F$ tests for model comparison for example. But Dr. Yu once said he never uses p-value in his own research, and what is taught in Intro Stats MATH 4720 is mostly problematic. In fact, the null hypothesis significance testing (NHST) paradigm and the p-value usage have been much criticized and shown to be problematic, misused, and resulting in reproducibility and replication crisis in scientific research. Please write a summary paper *at least two pages* including

  - Interpretation of p-value
  - List and discussion about the problems of the NHST and p-value method
  - Possible solutions to those problems
  
Some references are

  - [Wikipedia: Misuse of p-values](https://en.wikipedia.org/wiki/Misuse_of_p-values)
  - [A. Reinhart (2015), "Statistics Done Wrong", *No Starch Press*, San Francisco.](https://www.statisticsdonewrong.com/)
  - [R. L. Wasserstein and Nicole A. Lazar (2016), "The ASA Statement on p-Values: Context, Process, and Purpose", *The American Statistician*, 70:2, 129-133.](https://www.tandfonline.com/doi/full/10.1080/00031305.2016.1154108)
  - [V. Amrhein, S. Greenland and B. McShane (2019), "Retire statistical significance", *Nature*, 567, 305-307. ](https://www.nature.com/articles/d41586-019-00857-9)
  - [B. McShane, D. Gal, A. Gelman, C. Robert and Jennifer L. Tackett (2019), "Abandon Statistical Significance", *The American Statistician*, 73:sup1, 235-245.](https://www.tandfonline.com/doi/full/10.1080/00031305.2018.1527253)
  - [A. Gelamn and E. Loken (2014), "The Statistical Crisis in Science", *American Scientist*, 102, 460-465.](/http://stat.columbia.edu/~gelman/research/published/ForkingPaths.pdf)
  - [R. L. Wasserstein, A. L. Schirm and N. A. Lazar (2019), "Moving to a World Beyond $p<0.05$", *The American Statistician*, 73:sup1, 1-19.](https://www.tandfonline.com/doi/full/10.1080/00031305.2019.1583913)

  
There are lots of discussions and papers out there. You are welcome to google more resources to support your argument. The work should be entirely your effort. **You are not allowed to copy anyone's words, and you have to cite any resources you use, papers, blogs, videos, lecture notes, etc, or you violate Marquette academic misconduct policy.**

**Check last page**



# Statistical Computing and Data Analysis

Please perform a data analysis using $\texttt{R}$ or your preferred language. **Any results should be generated by computer outputs, and your work should be done entirely by a computer. Handwriting is not allowed. Relevant code should be attached.**


## Gasoline Mileage Data
We use the same data set `mpg.csv` for data analysis. For the following analysis, if any regressor contains a missing value `NA`, remove the corresponding row of the data matrix.
```{r}
mpgData <- read.csv("mpg.csv")
head(mpgData,3)
mpgData<-mpgData[c(-23,-25),]
```


1. Build a linear regression model relating gasoline mileage $y$ to vehicle weight $x_{10}$ and the type of transmission $x_{11}$ (1 automatic; 0 manual). Does the type of transmission significantly affect the mileage performance?

\quad\quad\ *Solution: *
```{r}

mpg_wt_t <- lm(y~x10+x11, data=mpgData)
summary(mpg_wt_t)
```
It does not do a great job, we would need to include some interaction term and test it again. The p-value is a little high, and the standard error is also too big compared with the estimate of $x_11$.

2. Modify the model developed in (1) to include an interaction between vehicle weight and the type of transmission. What conclusions can you draw about the effect of the type of transmission on gasoline mileage? Interpret the parameters in this model.

\quad\quad\ *Solution: *
```{r}
mpg_wt_t <- lm(y~x10*x11, data=mpgData)
summary(mpg_wt_t)
```
Now it is relevant, since there is a correlation between weight and type of transmission. The interaction term resolves the problem that we had before. The type of transmission seems to impact gas mileage, but the weight of the car highly correlated with the type of transmission and gas mileage. 


For the following questions (3) to (9), consider all the regressors except $x_4$, $x_5$ and $x_{11}$.

+ $y$: MPG
+ $x_1$: Displacement (cubic in.)
+ $x_2$: Horsepower (ft-lb)
+ $x_3$: Torque (ft-lb)
+ $x_6$: Carburetor (barrels)
+ $x_7$: No. of transmission speeds
+ $x_8$: Overall length (in.)
+ $x_9$: Width (in.)
+ $x_{10}$: Weight (lb)

*Normalize the response and predictors by unit length scaling* before later analysis.

3. Obtain the correlation matrix of regressors. Does it give any indication of collinearity? 

\quad\quad\ *Solution: *
```{r}

mpg_modif <- mpgData[,c(-5,-6,-12)]

mpg_scale <- mpg_modif

for(i in 1:length(mpg_modif))
{
  mpg_scale[,i]<-mpg_scale[,i]/norm(as.matrix(mpg_scale[,i]),"2")
}
X<- mpg_scale[,-1]
(XtX <- cor(X))

```

There are indications of collinearity between different parameters. 

4. Calculate the variance inflation factors (VIFs). Is there any evidence of collinearity?

\quad\quad\ *Solution: *
```{r}
full_mod <- lm(y~., data = data.frame(mpg_scale))
(vif_all <- car::vif(full_mod))
```

There is evidence of colinearity. $x_1, x_2, x_3, x_8, x_10$ all have variance inflation factors higher than 10, which is significant evidence of colinearity. 

5. Find the eigenvectors associated with the smallest eigenvalues of ${\bf X}'{\bf X}$. Interpret the elements of these vectors. What can you say about the source of collinearity in these data?

\quad\quad\ *Solution: *
```{r}
eigen_XtX <- eigen(XtX)
(lambda <- eigen_XtX$values)
```
Collinearity seems to come from $\lambda_6,\lambda_7,\lambda_8,\lambda_9$. However, that does not tell us which $x$, because $\lambda$s are ordered. 


6. Compute the condition indices and variance-decomposition proportions. What statements can you make about collinearity in these data?

\quad\quad\ *Solution: *
```{r}
max(lambda) / lambda
```
Only $k_8$ has a value taht is higher than 1000, which implicates that there is colinearity in that specificy lambda value. 


7. Fit the ordinary multiple regression and the ridge regression. Use the ridge trace to select an appropriate value of $\delta$. Explain how the ridge regression coefficients change as the parameter $\delta$ increases. This gives you the idea of why the ridge regression is a *shrinkage* method.

\quad\quad\ *Solution: *
```{r}
delta <- seq(0, 0.5, by = 0.01)
ridge_fit <- MASS::lm.ridge(y ~ .-1, 
                            data = data.frame(mpg_scale), 
                            lambda = delta)
matplot(coef(ridge_fit), type = "l",
        xlab = "delta", 
        ylab = "Coef", 
        main = "Ridge Trace")
abline(v = which(delta == 0.25), 
       col = "orange", lty = 2)
```

It is possible to see that all the lines are converging to almost the same result. The ridge regression coefficients change as the parameter $\delta$ increases, since we are introducing some bias at the cost of $R^2$, but after some time it will converge to a stabilized tradeof between the bias and $R^2$.


8. Use the all-possible-regressions approach to find an appropriate regression model.

\quad\quad\ *Solution: *
```{r}
olsrr_all <- olsrr::ols_step_all_possible(full_mod)
print(olsrr_all[which.max(olsrr_all$adjr),])
print(olsrr_all[which.max(olsrr_all$rsquare),])
print(olsrr_all[which.min(olsrr_all$cp),])
```

```{r}
olsrr::ols_step_best_subset(full_mod, metric = "adjr")
```
The appropriate model is given by using the predictors $x_7,x_8,x_10$

9. Use stepwise regression to specify a subset regression model. Does this lead to the same model found in (8)?

\quad\quad\ *Solution: *
```{r}
(olsrr_both <- olsrr::ols_step_both_aic(full_mod))
```

This does not lead to the same model, it suggests to only use  $x_1$ as a predictor.




